Are we learning to build better models at the cost of better decisions?

### 2021-03-05 manuscript/manuscript.Rmd
millie comments

@@

@@

Like many decision problems, the task of setting a sustainable harvest quota appears to hinge on having an accurate forecast: if we can predict to what size the fish stock will increase next year, $X_t+1$, and we know the current stock, $X_t$, then we can safely harvest $X_{t+1} - X_t$.  Overestimating or underestimating such recruitment will result in over-harvesting or under-harvesting, respectively.  Selecting a model based on forecast skill is also justifiable on theoretical grounds, since it reduces the risk of over-fitting by comparing model predictions to later observations that were not used to estimate the model. @Gneiting2007 provides a rigorous proof for "proper" scoring criteria, which have the desirable property which no model predicting the distribution of future outcomes, $Q(x)$ can achieve a better average score than the true model $P(x)$. That is, unlike likelihood or other goodness-of-fit scores, it is impossible to overfit when conditioning on a strictly proper score -- since no model model can out-perform the true model. Not that strictly proper scoring rules score _probabilistic forecasts_ and not just point predictions, favoring models which accurately reflect the uncertainty over those which under-estimate it.  These features have made proper scoring rules for probabilistic forecasts a successful and popular approach for addressing model uncertainty in many other areas [@Gneiting2014; @Raftery2016] and a promising tool for evaluation of ecological forecasts [@Dietze2018].

@@

# Discussion

@@

@@

<!-- move to appendix?  too confusing/too long?-->
Is this result an artifact of optimal control? The analysis has relied on optimal control theory to determine the optimal decision associated with each model or distribution of models.  The resulting policies are mathematically optimal under the assumptions given [see @Reed1979 for stochastic harvest problem in particular and @Marescot2013 for discussion of Markov Decision Problems more generally].  In most realistic decision problems, model complexity currently precludes the calculation of an optimal control solution at all ["the curse of dimensionality", e.g. @Polasky2011; @Marescot2013]. Under the assumptions of any of the models considered here, the optimal control solution leads to what is known as a "constant escapement" policy, which seeks to bring biomass to the target escapement level, $B_{MSY}$ as quickly as possible, $H_t = \min(X_t - B_{MSY}, 0)$.  If the assumptions are not met, a constant-escapement policy may not be optimal, as in the case when observations of the stock $X_t$ are not perfect [@pomdp-intro].  While constant escapement is used as a basis of management in some real world cases such as salmon fisheries, most marine fisheries are based on a target of "constant mortality", ($F_{MSY}$). Notably, a constant mortality policy, implemented as $H_t = F_{MSY} X_t$, would not perform better under Model 1 than Model 2.  Constant mortality is optimal only under deterministic growth at equilibrium (where the policy is the same under constant escapement).  Does that mean the result is an artifact of assuming constant escapement rather than constant mortality as the basis for management?  No, it does not.  It is just as easy to construct a model under which the constant mortality matches the constant mortality of the true model while being a poor predictor.  Moreover, using a non-optimal policy would always beg the question of whether the poor outcomes were merely the result of using a non-optimal policy rather than a consequence of model uncertainty.  Because the space of possible actions is smaller than that of possible models, we should expect most decision problems to have their own "Model 1," a model that would lead to correct actions while being a poor predictor of future outcomes.  The trouble is that we have no theory which would not reject such a model.

@@

@@

<!-- Omit this?  Does it beg an example? -->
<!--
There is at least one technique which might address the problem presented here.  Recently developed techniques based on model-free deep reinforcement learning train multi-layer neural networks to approximate the optimal control policy without assuming or attempting to model the underlying dynamic process at all.  Feedback is based on cumulative reward, not short-term reward or predictive accuracy.  
-->

>>>>

<!-- I think the introduction could have a clearer storyline and better introduce some of the ideas you bring up later on in the methods/results. A quick outline idea:

  - Paragraph1: Model-based forecasts have become a cornerstone of applied science in disciplines as diverse as atmospheric science [weather?], economics [stock markets?], and ecology [dietz]. Seeking to inform a variety of decisions, near-term forecasting in these diverse fields has  simultaneously been encouraged by policy-makers [CITE?] and enabled by the rapid and recent increase in data availability and computational methods leading to a rapid expansion of thier use (or something...)

   - Paragraph 2: Introduce ecological decision making under uncertainty, rapid global change, etc. (kind of like your current opening bit)

  - Paragraph 3: But do better forecast always lead to better decisions?  Because model-based forecasts frequently play an important role in decision-making,it is commonly assumed [e.g. @Walters1981; @Clark2001; @Dietze2018] that this approach for addressing model uncertainty will also improve decision-making outcomes. 
            - While the literature has largely addressed issues of overfittting that might lead forecasts to misguide decisions, proper scoring methods blah blah [...introduce this concept]. 
            - Beyond overfitted models,... [maybe add something about Value of Information literature here?]

  - Paragraph 4: Your current paragraph on "I illustrate..." (I'd add into that paragraph something about the fact that you explore adaptive management?)

A second idea on framing: this maybe is a to divergent from the main point (or leads you into a post-modernist narrative that you'd prefer to avoid...) but it seems like you could reframe the whole thing as an argument for embracing an inclusive process of resource management that acknowledges pluralism.  
  (1) Policy makers often think "scientists" hold knowledge of "an environmental reality"
  (2) Despite this not being true (scientist disagree on a lot of things), even when "science" points to a more likely reality, that might be off in ways that can fundementally misguide decisions. 
  (3) The problem is deeper than acknowledging a "rashamon set" of models or multiple possible descriptions of of reality. Sometimes what appears closer to reality is actually off in a way that is critical to the decision.
-->

<!-- Open on real-world issues, uncertainty, particularly model uncertainty as key theme.  -->

@@

<!-- I feel like parts of the below paragrph should be in the abstract (i.e. "I illustrate how the most promising techniques from statistical approaches and decision theoretic approaches to decision making under uncertainty would be applied to this simple problem, and demonstrate that it both cases they lead us away from the model that produces the most desirable decisions towards worse outcome") -->

@@

<!-- I think this subsection should be renamed. it makes it a bit hard to follow the structure of the larger section. Seems like something like: "Statistical approaches: forecasts and proper scores" idk, but something that more closely aligns with the above paragraph and the second subsection -->

Like many decision problems, the task of setting a sustainable harvest quota appears to hinge on having an accurate forecast: if we can predict to what size the fish stock will increase next year, $X_t+1$, and we know the current stock, $X_t$, then we can safely harvest $X_{t+1} - X_t$.

<!-- it is unclear what you mean by safely harvest?  as in the stock wont crash? or it won't be worse than it is this year? You could imaging that we could safetly harvest more (or only less) than that.-->

Overestimating or underestimating such recruitment will result in over-harvesting or under-harvesting, respectively.

Selecting a model based on forecast skill is also justifiable on theoretical grounds, since it reduces the risk of over-fitting by comparing model predictions to later observations that were not used to estimate the model.

@Gneiting2007 provides a rigorous proof for "proper" scoring criteria, which have the desirable property which no model predicting the distribution of future outcomes, $Q(x)$ can achieve a better average score than the true model $P(x)$. That is, unlike likelihood or other goodness-of-fit scores, it is impossible to overfit when conditioning on a strictly proper score -- since no model model can out-perform the true model. Not that strictly proper scoring rules score _probabilistic forecasts_ and not just point predictions, favoring models which accurately reflect the uncertainty over those which under-estimate it.  These features have made proper scoring rules for probabilistic forecasts a successful and popular approach for addressing model uncertainty in many other areas [@Gneiting2014; @Raftery2016] and a promising tool for evaluation of ecological forecasts [@Dietze2018].

<!-- This scoring bit comes abruptly (and it feels like could be in the introduction) -->

@@

I think it would be effec

# Discussion
<!--
hit bigger themes first & objections first
- Learning can be counter-productive
- Value of information can be negative in practice

Objections:
- we already know simple models can be better, overfit models can lead us astray
- we already know Knightian uncertainty matters.  (underestimating uncertainty isn't the same thing)

@@

<!-- (3) THIS problem (4) THIS problem. I like that you outline the points you are going to cover! -->

@@

<!-- move to appendix?  too confusing/too long?

Not sure - it feels a little extraneous (I did think this was an interesting point brought up by reviewer 1 but I also don't necessarily think it warrants a long paragraph that might confuse the main point rather than clarify something) Could this be a couple sentences and reference an appendix that walks through what you write out (potentially even with a simple constant mortality example)
-->
Is this result an artifact of optimal control? The analysis has relied on optimal control theory to determine the optimal decision associated with each model or distribution of models.  The resulting policies are mathematically optimal under the assumptions given [see @Reed1979 for stochastic harvest problem in particular and @Marescot2013 for discussion of Markov Decision Problems more generally].  In most realistic decision problems, model complexity currently precludes the calculation of an optimal control solution at all ["the curse of dimensionality", e.g. @Polasky2011; @Marescot2013]. Under the assumptions of any of the models considered here, the optimal control solution leads to what is known as a "constant escapement" policy, which seeks to bring biomass to the target escapement level, $B_{MSY}$ as quickly as possible, $H_t = \min(X_t - B_{MSY}, 0)$.  If the assumptions are not met, a constant-escapement policy may not be optimal, as in the case when observations of the stock $X_t$ are not perfect [@pomdp-intro].  While constant escapement is used as a basis of management in some real world cases such as salmon fisheries, most marine fisheries are based on a target of "constant mortality", ($F_{MSY}$). Notably, a constant mortality policy, implemented as $H_t = F_{MSY} X_t$, would not perform better under Model 1 than Model 2.  Constant mortality is optimal only under deterministic growth at equilibrium (where the policy is the same under constant escapement).  Does that mean the result is an artifact of assuming constant escapement rather than constant mortality as the basis for management?  No, it does not.  It is just as easy to construct a model under which the constant mortality matches the constant mortality of the true model while being a poor predictor.  Moreover, using a non-optimal policy would always beg the question of whether the poor outcomes were merely the result of using a non-optimal policy rather than a consequence of model uncertainty.  Because the space of possible actions is smaller than that of possible models, we should expect most decision problems to have their own "Model 1," a model that would lead to correct actions while being a poor predictor of future outcomes.  The trouble is that we have no theory which would not reject such a model.

@@

<!-- I really like the last paragraph!! -->

@@

<!-- Omit this?  Does it beg an example?

There is at least one technique which might address the problem presented here.  Recently developed techniques based on model-free deep reinforcement learning train multi-layer neural networks to approximate the optimal control policy without assuming or attempting to model the underlying dynamic process at all.  Feedback is based on cumulative reward, not short-term reward or predictive accuracy.  
-->

<!--I was thinking about exactly this while I was reading the paper.  I do think this would likely beg an example, which might be a can of worms that you don't want to open with this piece... I also don't think that I think it would address the presented problem. (Kind of repeating my second framing point but) It reminds me of the "Stop Explaining Black Box Machine Learning Models for High Stakes Decisions and use Interpretable Models Instead" paper where she talks about how, for a given data set, you can define the Rashomon set as the set of reasonably accurate predictive models-  Because the data are finite, the data could admit many close-to-optimal models that predict differently from each other: a large Rashomon set. I suspect this happens often in practice because sometimes many different ML algorithms perform similarly on the same data set, despite having different functional forms (for example, random forests, neural networks, support vector machines). As long as the Rashomon set contains a large enough set of models with diverse predictions, it probably contains functions that can be approximated well by simpler functions, and so the Rashomon set can also contain these simpler functions. Said another way, uncertainty arising from the data leads to a Rashomon set; a larger Rashomon set probably contains interpretable models, thus interpretable accurate models often exist. If this theory holds, we should expect to see interpretable models exist across domains. These interpretable models may be hard to find through optimization, but at least there is a reason we might expect that such models exist.

The problem is deeper than acknowleding a "rashamon set of models" or a plurality of reality (relevant to that Bioscience paper I sent last week about the Rashamon effect in conservation). Sometimes what appears closer to reality is actually fundementally wrong in some way that is misguiding to a decision maker...more opaque approaches could lessen our capacity to trouble shoot that?) -->
